---
title: " Introduction to Multisplit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multisplit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction

The `Multisplit` package provides tools for splitting datasets
with multiple binary outcomes into training and test sets.

## Splitting a dataset

First, we create a dataset containing multiple binary outcomes.
```{r}
library(Multisplit)

data <- data.frame(
  outcome1 = rep(c(0, 1), 10),
  outcome2 = rep(c(0, 0, 1, 1), 5)
)

head(data)
```
We can then use "split_multioutcome()" to divide the data
into training and test sets.
```{r}
result <- split_multioutcome(
  data = data,
  outcomes = c("outcome1", "outcome2"),
  k = 10,
  train_folds = 7,
  seed = 1234
)
```

##Training and test sets

The resulting object contains the training set, the test set,
and a summary table of the distribution of positive cases.
```{r}
train_set <- result$train
test_set <- result$test
balanced_table <- result$balanced_table
```
The number of observations in each set can be checked with:
```{r}
nrow(train_set)
nrow(test_set)
```
The balanced_table can be used to assess the distribution
of positive cases across the datasets:
```{r}
balanced_table
```

##Conclusion

The "split_multioutcome()" function provides a flexible approach
for splitting datasets with multiple binary outcomes while
attempting to maintain a balanced distribution of positive cases.
